% Kings County 2.5 yr peated bourbon (Brooklyn, NY)

\chapter{ANALYSIS}
\label{ch:analysis}

The angular acceptance of St.\ George can be determined by comparing the
expected yield from the reaction at the desired energies to the actual
counts measured at the detector plane. If we assume a symmetric angular
acceptance, the opening angle of the acceptance cone can be directly
calculated and compared to the anticipated angular acceptance opening
angle of 40~mrad. The angular acceptance and its uncertainty for each
energy can be determined independently of the other points. The angular
acceptance is based on the properties of the target, incident beam,
detector, and recoil separator and the decisions relating to the
experiment itself.

The measurement of the yields near and on the two resonances can also be used
to determine the properties, such as the resonance strength, related to those
resonances.

All code was written using the scientific Python stack (numpy, scipy,
matplotlib, pandas, pymc3). An analysis framework pyne was developed
concurrently in python to aide in direct analysis of the run information
from the raw data files.


\section{Target Properties}
\label{sec:target-properties}

A self-supporting \nuc{27}{Al} target was used for the entirety of the
experiment. The target thickness was measured using an offline detector
station and a mixed \nuc{241}{Am}/\nuc{148}{Gd} alpha-particle source. The
target thickness was measured by observing the energy loss of the two
alpha peaks as compared to the direct detection of the alpha particles
by the detector. That energy loss is converted to a target thickness in
$\mu$g/cm${}^{2}$ which can be used to determine the energy loss for other
particles at any energy. The measured thickness was $63.8^{+2.0}_{-1.7}$~
$\mu$g/cm${}^{2}$.

% see foil_thickness/target_thickness_20180403.ipynb

Two runs were performed, one without the target in place and one with
the target in place, using an annular Si detector placed within a vacuum
chamber. Each run lasted approximately 600~s and were performed
immediately following each other. The detector was calibrated using the
spectrum from the first run without the target in place, using the known
energies of the peaks (see Table XX). The energy calibration parameters
were determined through a combined Monte Carlo and Bayesian procedure, as
outlined below.

\textbf{Step 1:}
  New spectra were simulated by sampling a single value for each bin
  from the Poisson distribution $n \sim \rm{Poisson}(\lambda)$,
  where $n$ is the simulated counts in that bin and $\lambda$ is the scale
  parameter for the distribution, which is taken to be the detected
  counts for that bin. A total of 2000 spectra were generated in this
  fashion.

\textbf{Step 2:}
  The two largest peaks were identified first through a continuous
  wavelet transformation to get an approximate peak center. The final
  peak center was determined by taking a weighted average of the eleven
  bins centered on the selected bin to get the final bin value for the
  peak center.

\textbf{Step 3:}
  A linear regression is fit, using the peaks found from the simulations
  and the known calibration energies of the alpha peaks. By fitting the
  entire sample of peak locations, we determine a global fit for our
  detector. The linear model used is a Bayesian model described by:

\begin{align*}
    E &\sim \rm{Normal}(\mu, \sigma) \\
    \mu &= \alpha + \beta \rm{bin} \\
    \alpha &\sim \rm{Normal}(28, 10) \\
    \beta &\sim \rm{Normal}(5, 5) \\
    \sigma &\sim \rm{Lognormal}(2, 10)
\end{align*}

  The Bayesian model allows for the uncertainty on the fit to be
  propagated forward.

Table XX: Alpha-particle energies for the 241Am/148Gd mixed source:

\begin{verbatim}
148Gd   3182.69     100
241Am   5388          1.6
        5442.8       13.1
        5485.56      84.8
\end{verbatim}

From the linear model, an energy calibration can be applied to the
in-target run to determine the energy of the two largest peaks. Due to
the lower count rate due to the target being in place, only the largest
peaks could be identified. The procedure to determine the energy loss is
similar to that previously described for the energy calibration:

\textbf{Step 1:}
  New spectra based on the spectrum with the target in place were
  generated in the same manner as before. A total of 2000 spectra were
  generated.

\textbf{Step 2:}
  For each spectra, a linear model was sampled from the fitted Bayesian
  linear model and applied to the spectra to convert the bin numbers to
  energy.

\textbf{Step 3:}
  The two largest peaks were identified first with a continuous wavelet
  transformation then a weighted average of the counts to find the central
  energy of the peak.

\textbf{Step 4:}
  The energy of the peaks were subtracted from the known energies of the
  peaks to determine the energy loss through the sample for two known
  input energies. The uncertainty in the sample alpha energies was
  assumed to be negligible and was not considered in the calculations.

This energy loss may be related to the target thickness through the use
of the computer code SRIM. The stopping and range table for alpha
particles passing through \nuc{27}{Al} was generated. A linear spline
(i.e. each data point from the range table connected by a linear fit)
was generated for the range, and the range for each fitted peak energy
from the spectra generated from the in-target run was determined. This
range was subtracted from the alpha ranges for the known calibration
energies and multiplied by the density of aluminium to getthe thickness
in $\mu$g/cm${}^{2}$. The uncertainties in the calibration alpha energies and
in the SRIM data were assumed to be negligible and were not considered
during the analysis.

Finally, as each simulated spectra was assumed to be an independent
sample, the two thicknesses determined from the two alpha peaks were
averaged to get a single value for the thickness from each iteration.
The thickness distribution generated in this case was compared against
the dsitribution of 4000 samples (2 dependent thickness measurements for
each of the 2000 samples), with the averaged thickness case exhibiting a
slightly narrower distribution. The averaged thickness samples were used
for the remainder of the analysis. The thickness of $63.8^{+2.0}_{-1.7}$~
$\mu$g/cm${}^{2}$ is the mean and 95\, \% confidence interval from this
distribution of thickness samples. For subsequent calculations, the
thickness used was sampled from this distribution.

The uncertainties present in each step of the procedure laid out above
are automatically propagated forward due to the methods chosen. The
stoichastic nature of the process allows the influence of the base
assumptions of the underlying data (e.g. the counts in each bin are
drawn from a Poisson distribution) to be seemlessly brought forward
without the need of clumbersome mathematics that can potentially hide
wrong assumptions about the values in question, such as that all of the
data is normally distributed.

For most of the following calculations, the number of target nuclei per
square centimeter is used instead of the thickness in $\mu$g/cm${}^{2}$. Our
target thickness is then $1.42^{+0.06}_{-0.06} \times 10^{18}$ nuclei/cm${}^{2}$.
This value is useful for calculating the energy loss of the proton
through the target, as that relies on the number density of the target
and the stopping power of the material. The energy loss of the beam will
be discussed in the next section.


\section{Beam Properties}
\label{sec:beam-properties}

The incident proton beam was produced by the 5U and delivered to the St.
George target area. The beam energy and resolution were determined
through a series of accelerator and beamline commissioning experiments
performed before this experiment was performed.

% Similar experiments showed that the energy resolution
% of the beam is approximately 300 keV. To be conservative, a value of 500 keV
% was used for all runs.

The beam energy was determined from the calibration of the 5U analyzing
magnet performed during a different experiment. During the experiment,
the magentic changes were performed slowly such that the magnetic field
did not appreciably drift during the runs. The energy resolution can
also be determined from the calibration runs, where the resolution is
given by the energy width of the leading edge of the resonance scan.
Values of approximately 300 eV were commonly observed, with a
conservative value of 500 eV adopted for this experiment since no direct
energy calibration was performed with our specific experimental setup.
The uncertainty in the analyzing magnet field is accounted for within
this uncertainty and is not considered separately.

% While the uncertainty in the field strength recorded is
% not completely negligible, it is included within the beam energy resolution
% and is not considered separately.

The beam current was relatively stable during the experiment. During the
longer runs, the beam current was measured every 15 minutes in order to
monitor its change during the run. For each run, the current uncertainty
was determined by the measured values for cases where multiple current
measurements were performed, or 5\,\%. For all runs, the final current
uncertainty was between 5 and 12\,\%. Ideally, an offset Si detector at
the target location would be used to monitor the beam current during the
entirety of the run by measuring the current of the scattered beam
particles at a fixed angle. As this setup was not available for the
target chamber, periodic direct measurements of the current using the
Faraday cup at the target location were required to measure the beam
intensity.

\section{Detector Properties}
\label{sec:detector-properties}

A 16-strip Si detector was used to detect the produced alpha particles
during the experiment. A calibration run was performed following the
experiment using the same detector and data acquisition settings as used
during the experiment. A \nuc{241}{Am}/\nuc{148}{Gd} mixed alpha source was used
for calibrating the energy conversion and energy resolution of each
strip separately. All of the strips were similar with approximately 2
keV/bin for the calibration and approximately 2.75\,\% (90~keV) for the
energy resolution.

The calibration run resulted in a single spectrum. Due to the poor
energy resolution of the detector resulting from the lower-than-optimal
bias voltage setting used during the experiment, only the two highest
intensity peaks could be resolved above the background. As the alpha
peak resulting from \nuc{148}{Gd} is closer in energy to the alpha
particles produced in the experiment, The alpha peaks also exhibit long
low-energy tails such that the particles produced in the reaction are
smeared out in energy. For the experimental run, an energy threshold was
set to exclude incident proton counts, where counts appearing above the
threshold are considered to be from alpha particles. That threshold was
set by the following:

\begin{equation}
    E_{\rm proton} + 3 \sigma_{\rm beam} + 3 \sigma_{\rm resolution}
\end{equation}

The detector efficiency was not directly measured and assumed to be
100\,\%. Efficiency measurements performed during the commissioning work
supporting the St.\ George detector system resulted in efficiencies above
99\,\% for all strips.

A simulation of the expected energy spectrum at the target location was
performed using SRIM data tables. The simulation looked at the known
energy loss within the target of the incident beam, the expected cross
section within the energy limits of the target, the energy loss of the
produced alpha particles through the remainder of the target, and the
energy resolution of the detector to generate an expected energy
spectrum. The procedure for this simulation is as follows:


\textbf{Step 1:}
  An energy deviation drawn from the Normal(0, sigma) distribution
  (where sigma is the beam energy resolution) for 2000 particles. This
  energy deviation is the difference in energy from the central energy.

\textbf{Step 2:}
  SRIM files for the central energy were generated for fractional depths
  within the target, where the output is the beam energy profile at that
  target depth.

\textbf{Step 3:}
  Using the expected cross section from the AZURE2 R-matrix calculation,
  a depth for each of the simulated particles was generated to determine
  the location within the target that the reaction takes place.

\textbf{Step 4:}
  A beam energy $E_d$ is generated from the distribution of beam energies
  at the given depth, and the initial deviation for that particle is
  added to the energy to give the final beam energy.

\textbf{Step 5:}
  The beam energy is converted to the produced alpha particle energy
  through the kinematic equation

\begin{equation}
    E_{\alpha} = Q + E_{\rm p}\left(1 - \frac{m_{\rm p}}{m_{\rm p} + m_{\nuc{27}{Al}}}\right)
\end{equation}


\textbf{Step 6:}
  The deviation of the alpha particle from a known alpha energy (used to
  generate SRIM files at various depths) was recorded.

\textbf{Step 7:}
  An alpha energy was generated for each particle based on the remainder
  of the target that it needs to travel through from the final energy
  distribution generated for particles traveling through that thickness.

\textbf{Step 8:}
  The energy deviation is added back to the alpha particle's energy to
  give its final energy.

This procedure generates an alpha-particle energy spectrum following the
target location given the known parameters about the target thickness
and the cross section, and incorporates the known energy resolution of
the incident alpha beam and the stoichastic nature of the energy loss
and reaction within the target. Finally, using the known energy
resolution of the detector, a final energy spectrum can be generated by
drawing new alpha particle energies from the distribution:

\begin{equation}
    E_{\alpha,\rm{detector}} \sim \rm{Normal}(E_{\alpha},\sigma_{\rm detector})
\end{equation}

An example of the output of this procedure is given in {[}FIGURE{]},
where the agreement between the location and width of the alpha peak can
be seen in the normalized spectra. Note that the low energy tailing of
the detected particles is not modeled in our simulated spectrum, as we
don't know the full characteristics for the detector response.


\section{Additional Parameters}
\label{sec:additional-parameters}

Additional inputs into the final calculation of the acceptance of St.
George are the cross section determined from an R-matrix fit on several
low-lying resonances, the stopping power of protons in aluminium from
SRIM, the run time, and the counts at the detector. For those parameters
that are derived from external programs (AZURE2 and SRIM), the
uncertainty is assumed to be negligible. The uncertainty in the time was
assumed to be 10 seconds for those runs that only lasted for a single
15-minute span, and higher for those runs that required the periodic
measurement of the beam current which resulted in stopping the incident
beam for an unspecified duration of time.

The counts at each detector were the sum of all events above the
threshold defined by the beam energy and detector resolution. The counts
are Poisson distributed, with the length of time for the run was such
that the uncertainty from the counts at the detector was not above 5\,\%,
with most runs having a count uncertainty of a much lower value. The
direct uncertainty of the counts at the detector is partially convolved
with the run time; a lower counting uncertainty requires a longer run
time and potentially a larger time uncertainty.

The direct beam reduction by St.\ George must be on the order of
$10^{10}-10^{14}$ in order to avoid damaging the Si detector and to
measure lower value regions of the cross section. This requirement is
within the designed capabilities of St.\ George when tuned for heavy
recoil transmission to the final detector plane, but must be verified
experimentally due to the altered tune and different detector plane
required for this experiment. During the experiment, count rates at the
detector were monitored, and potential counts from the direct proton
beam were excluded from the final counts with the energy discriminator
previously described.

\section{Final Acceptance Measurements}
\label{sec:final-acceptance-measurements}

The acceptance of St.\ George can be found for each energy value by
comparing the detected counts to the expected yield for that incident
beam energy. The yield is found from:

\begin{equation}
    Y(E) = \frac{N_r}{N_b},
\end{equation}

where $N_r$ is the number of reaction products produced and $N_b$ is the
number of incident beam particles. We can determine $N_b$ from the beam
current and the total run time. The value for $N_r$ is determined by the
total counts at the detector (for the experimental yield) or the
integration of target and cross section properties following

\begin{equation}
    Y(E_0) = \int \frac{\sigma(E)}{\epsilon(E)}\,\rm{d}E
\end{equation}

In both cases, the detector efficiency and St.\ George transport
efficiency are 100\,\%, as previously discussed.

The acceptance in mrad is given by

\begin{equation}
    \theta = \arccos\left(1 - 2 \frac{Y_{\rm experiment}}{Y_{\rm theory}}\right)
\end{equation}

The angular acceptance can be calculated in this way for each run
individually, as shown in FIGURE and TABLE. The process for
calculating the uncertainty bounds is given by the following, repeated
2000 times to have enough confidence in the final values:

\textbf{Step 1:}
  The beam energy, beam current, and time are sampled from a normal
  distribution $\rm{Normal}(\mu,\sigma)$, where $\mu$ and $\sigma$ are for the value
  (energy, current, or time) in question.

\textbf{Step 2:}
  The incident number of particles is calculated from the current and
  time.

\textbf{Step 3:}
  The target thickness in energy is calculated from finding the stopping
  power at the incident beam energy from the SRIM tables, and sampling
  from the distribution of target thicknesses in terms of atoms/cm${}^{2}$.

\textbf{Step 4:}
  The yield is determined by integrating EQUATION between the
  entrance energy and the lower energy given by that entrance energy
  minus the target thickness.

\textbf{Step 5:}
  The experimental yield is drawn from a poisson distribution
  $\rm{Poisson}(c)$, where $c$ is the number of counts detected.

\textbf{Step 6:}
  The acceptance for the iteration is caluclated by EQUATION.

The distribution of values generated by the process above can be used to
find the acceptance and confidence intervals for the run in question.
Each run has an acceptance described by its distribution, which is the
run for that particular setting of St.\ George.


\section{Resonance Parameters}

Determine resonance parameters

Find area under yield curve

Relate to equation

Repeat a bunch of times to include energy, yield, and acceptance uncertainty (?)

Two to find: resonance strengths, reaction rates
